I don't know, are you sure we want to do this?\\ Sea level adaptation decisions under uncertainty

### 2016-11-09 submission/paper.tex
Minor edit in sec 5
Add code for Bergen decision-making
updates to Bergen example
Added  first stab at key terms

\begin{keypoints}
\item Keypoint 1
\item Keypoint 2
\item Keypoint 3
\end{keypoints}

@@

Figure \ref{fig:bergenobs} shows uncorrected and corrected Bergen sea level data, and the relationship between the corrected Bergen data and the global sea level data. The glacial isostatic adjustment is 0.26 (0.07) cm/yr. The time series regression uses an ARMA(1,1)-model \citep{boxjenkins}, with AR parameter 0.82 (0.13), and MA parameter --0.61 (0.17). The regression slope is 1.30 (0.12).
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{bergenfit.png}
\caption{ The left figure shows raw (black)and gia-corrected (red) sea level data from Bergen, The right figure relates the gia-corrected Bergen sea level to the global sea level series of \citet{csiro}. The straight line is the time series regression line.}
\label{fig:bergenobs}
\end{center}
\end{figure}

For the relationship between global annual mean temperature and global annual mean sea level rise we use the results from \citet{Bolin2014a}. The left panel of figure \ref{fig:ci} shows the simultaneous 90 \% confidence region for Bergen sea level rise relative to 1999 under scenario RCP 8.5, which is the scenario Norwegian authorities recommend for planning purposes.
\begin{figure}
\begin{center}
\begin{minipage}{.5\textwidth}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{DamageDistribution.png}
\caption{ The estimated distribution of annual storm surge damage in Bergen for 2015 (red) based on observed annual damage in Hordaland and Rogaland 1980-2015 (gray bars). }
\label{fig:BergenDamageDist}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{EuropeanIncreaseLossExtrapolationUncertainty.png}
\caption{ Relative change in mean annual damage as a function of sea level rise for 15 European cites as estimated by \cite{Hallegatte&2013} (black circles) with linearly extrapolated values indicated by gray lines. The median change and the corresponding extrapolation are indicated in red.}
\label{fig:ChangeInDamage}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{AccumulatedFutureLoss.png}
\caption{Distribution of log  accumulated future loss 2016-2100 under RCP 8.5 and assuming that no adaptation measures are implemented.} 
\label{fig:NoAction}
\end{center}
\end{figure}

@@

In many cases sea level rise projections are given as a single number for each scenario, usually the mean or median of the ensemble of projections from different climate models (e.g. \citet{climateimpactgroup}). Sometimes the spread of the ensemble is used to assess the uncertainty in the projections (e.g., the Norwegian Environmental Agency recommends using the upper ensemble value for RCP 8.5 as the basis for planning decision, pers. comm. from Even Nilsson, Norwegian Mapping Authority). In our analysiis there are two more sources of uncertainty, namely the two regression models. Figure \ref{fig:unc} shows the single number (vertical black line), the ensemble spread (histogram), the uncertainty including the global model (red) and the full uncertainty (blue) for Bergen projections of sea level rise relative to 1999 under RCP 8.5.

\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{unc.pdf}
\caption{2050 Bergen sea level projections with uncertainty due to different sources for RCP 8.5. The black line is the median projection (with no uncertainty), while the grey histogram corresponds to the spread of the climate models, the red curve adds the uncertainty due to the relation between global temperature and global sea level, and the blue line that due to downscaling global sea level to Bergen. } 
\label{fig:unc}
\end{center}
\end{figure}

>>>>

\begin{keypoints}
\item Decisions on adaptation measures need to take careful account of uncertanties
\item More careful models are needed for increased costs of the effects of sea level rise
\item Modelling local sea level rise is essential
\end{keypoints}

@@

Figure \ref{fig:bergenobs} shows uncorrected and corrected Bergen sea level data, and the relationship between the corrected Bergen data and the global sea level data. The glacial isostatic adjustment is 0.26 (0.07) cm/yr. The time series regression uses an ARMA(1,1)-model \citep{boxjenkins}, with AR parameter 0.82 (0.13), and MA parameter --0.61 (0.17). The regression slope is 1.30 (0.12).
\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=\linewidth]{bergenfit.png}
\caption{ The left figure shows raw (black)and gia-corrected (red) sea level data from Bergen, The right figure relates the gia-corrected Bergen sea level to the global sea level series of \citet{csiro}. The straight line is the time series regression line.}
\label{fig:bergenobs}
\end{center}
\end{figure}

For the relationship between global annual mean temperature and global annual mean sea level rise we use the results from \citet{Bolin2014a}. The left panel of figure \ref{fig:ci} shows the simultaneous 90 \% confidence region for Bergen sea level rise relative to 1999 under scenario RCP 8.5, which is the scenario Norwegian authorities recommend for planning purposes.
\begin{figure}[!hbpt]
\begin{center}
\begin{minipage}{.5\textwidth}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.5\linewidth]{DamageDistribution.pdf}
\caption{ The estimated distribution of annual storm surge damage in Bergen for 2015 (red) based on observed annual damage in Hordaland and Rogaland 1980-2015 (gray bars). }
\label{fig:BergenDamageDist}
\end{center}
\end{figure}

@@

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=0.5\linewidth]{EuropeanIncreaseLossExtrapolationUncertainty.pdf}
\caption{ Relative change in mean annual damage as a function of sea level rise for 15 European cites as estimated by \cite{Hallegatte&2013} (black circles) with linearly extrapolated values indicated by gray lines. The median change and the corresponding extrapolation are indicated in red.}
\label{fig:ChangeInDamage}
\end{center}
\end{figure}

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=\linewidth]{AccumulatedFutureLoss.png}
\caption{Distribution of log  accumulated future loss 2016-2100 under RCP 8.5 and assuming that no adaptation measures are implemented.} 
\label{fig:NoAction}
\end{center}
\end{figure}

@@

In many cases sea level rise projections are given as a single number for each scenario, usually the mean or median of the ensemble of projections from different climate models (e.g. \citet{climateimpactgroup}). Sometimes the spread of the ensemble is used to assess the uncertainty in the projections (e.g., the Norwegian Environmental Agency recommends using the upper ensemble value for RCP 8.5 as the basis for planning decision, pers. comm. from Even Nilsson, Norwegian Mapping Authority). In our analysiis there are two more sources of uncertainty, namely the two regression models. Figure \ref{fig:unc} shows the single number (vertical black line), the ensemble spread (histogram), the uncertainty including the global model (red) and the full uncertainty (blue) for Bergen projections of sea level rise relative to 1999 under RCP 8.5. We see that the ensemble range is about 16 cm, whereas the overall uncertainty range is about 40 cm.

\begin{figure}[!hbpt]
\begin{center}
\includegraphics[width=\linewidth]{unc.pdf}
\caption{2050 Bergen sea level projections with uncertainty due to different sources for RCP 8.5. The black line is the median projection (with no uncertainty), while the grey histogram corresponds to the spread of the climate models, the red curve adds the uncertainty due to the relation between global temperature and global sea level, and the blue line that due to downscaling global sea level to Bergen. } 
\label{fig:unc}
\end{center}
\end{figure}
